Title

### 2018-09-18 thesis/paper/paper.tex
Typos

While this is similar to RDF and Wikidata has in fact been exported as RDF \cite{EGKMV2014}, the data structure is more complex than simple triples. It allows to store additional information about statements. This is done via a statement ID, which is assigned to every triple storing information about an item. Statements have one of three ranks, \emph{normal}, \emph{preferred} and \emph{deprecated} which are used to differentiate similar statements on the same item. For example, the \emph{preferred} rank would be given to the current \emph{head of state} (P35) of a country, while all statements denoting former \emph{heads of state} have the rank \emph{normal}. \emph{Deprecated} is used on statements that where once considered correct but are not anymore, e.g., \emph{Chelsea Manning} (Q298423) having the \emph{sex or gender} (P21) \emph{male} (Q6581097).

Every statement can have added qualifiers, which consist of the corresponding statement id, a predicate and an object. Qualifiers provide contextual information for statements. The most common context is the time frame in which the statement was accurate, e.g. stating the \emph{start time} (P580) and \emph{end time} (P582) of the term of a head of state, but other information is possible, e.g. determination method for measurements or the elevation above sea level for the highest point. Additionally, references can be added to statements.

References support the claim by citing the source and additional information, e.g. when it was retrieved or when the source was published as property-value pairs. Each references consists from multiple such pairs, and statements can have multiple references.

@@

\section{Datalog}
Datalog is a logic query language. It is fully declarative and uses rules to derive additional knowledge from given facts. The most basic domain element in datalog is a \emph{constant}, which will be denoted as \emph{someConstant}. In the context of Wikidata, any item or property, e.g. \emph{Earth} (Q2) or \emph{head of government} (P6), could be represented as a constant. Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\), written as predicate\([n]\). To stay close to Wikidata in our examples let us take the predicate triple\([3]\) as an example. This would be used to express the \emph{item}-\emph{property}-\emph{value} relation with each position representing one part of the triple.\\
Additionally, datalog also provides \emph{variables}, which will be capitalized like \(X\). The sets of predicates {\bf P}, constants {\bf C} and variables {\bf V} are mutually disjoint. Constants and variables are collectively known as \emph{terms} and are, together with predicates, used to form \emph{atoms}. An atom has the form predicate\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog. A rule is a premise that implies a conclusion, and both are conjunctions of atoms.
Let \(P_n\) denote  premise atoms and \(C_n\) the conclusion atoms. The variables used in one rule are divided into three mutually disjoint sets \(V_p\), \(V_c\), and \(V_r\) which are the variables occurring only in its premise atoms, only in its conclusion atoms, and occurring in both, respectively. The general from of an existential rule is (with \(k \in \mathbb{N}^+\), \(l \in \mathbb{N}\))
\begin{equation}
\forall V_p \forall V_r.(\exists V_c.C_1 \wedge \ldots \wedge C_k \leftarrow P_1 \wedge \ldots \wedge P_l)
\end{equation}

\(V_c\) are called existential variables. If \(V_c = \emptyset\) the rule is called a \emph{datalog rule} and if \(l = 0\) it is called a \emph{fact}. If a fact contains no variables but only constants it is a \emph{ground rule}. Since all \(V_p\) and \(V_r\) are  quantified, the preceding universal quantifiers \(\forall V_p \forall V_r\) will be omitted for the rules presented in this work.

@@

\begin{example}
For the following example, taken from Wikidata, descriptive names will be used instead of the IDs to increase readability. The individual entities and properties are introduced with their ID beforehand so the example can be retraced. The species \emph{blue whale} (Q42196) has the \emph{parent taxon} (P171) \emph{balaenoptera} (Q133320), which in turn has the parent taxon \emph{rorqual} (Q232829). The predicates used are triple\_EDB\([3]\) as the input and isInTaxon\_IDB\([2]\), which will be derived. The basic facts are:
\[\text{triple\_EDB}(\text{blue whale}, \text{parent taxon}, \text{balaenoptera})\]
\[\text{triple\_EDB}(\text{balaenoptera}, \text{parent taxon}, \text{rorqual})\]
From this we can draw conclusions. The first step is to represent the already existing facts as IDB:
\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{triple\_EDB}(X, \text{parent taxon}, Y)\]
Finally the transitive closure of this relation can be derived with
\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{isInTaxon\_IDB}(X, Z) \wedge \text{isInTaxon\_IDB}(Z, Y)\]
\end{example}

@@

To improve the data quality, to help editors avoid common mistakes and to clarify the usage of properties, Wikidata allows constraints to be defined on properties. They are, however, not firm rules and exceptions can and should be made if necessary. A property constraint says something about the way a property may or may not be used. This can mean simple restrictions on the statement a property is part of, e.g. that only certain values are allowed, or more complex requirements on connected parts of the data, e.g. that the value of a statement with this property should have certain statements. In Table \ref{tab_property_constraints} you can find the existing property constraints as of September 2018. Afterwards the constraints will be explained in detail and their modelling given. For constraint that cannot be modelled, reasoning for this assessment is included.

To apply a constraint to a property, a statement <property> <P2302> <constraint> is added to the properties item page, where <property> is the property that should be constrained, <P2302> (property constraint) denotes that the property has a constraint and <constraint> is the specific constraint. Note that a constrained statement means a statement whose predicate is a property that has a specific constraint. A simple example for this would be \emph{distinct values} (Q21502410) stating that values of the property must be unique across all constrained statements. For this case no additional information regarding the constraint is necessary and violation could be found by surveying all constrained statements and comparing their values.\\
Constraints can be further specified by using qualifiers on the constraint statement. These qualifiers are either optional or mandatory and may be applied once or multiple times. For example, the \emph{none of} constraint (Q52558054) states that values for the specific property may not be in a given set of disallowed values. This set is specified using qualifiers with property \emph{item of} (P2305), which list all the disallowed values. At least one such qualifier must be present for the statement.\\
More complex dependencies are also possible. A constraint can depend on all constrained statements of a specific item (\emph{single value} constraint [Q19474404]), on all statements of the item specified as the value of a constrained statement (\emph{value requires statement} constraint [Q21510864]) or on all statements using the subclassOf-predicate (\emph{type} constraint [Q21503250]).

Table \ref{tab_property_constraints} shows all constraints that are used as of September 2018. It is roughly sorted by how difficult it is to model the given constraint using logical rules, based on the required expressive features and pre-processing. These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing of the dataset. Four constraints do not need any pre-processing except for loading the data into datalog as ground rules. Six can be solved if constant inequality is defined. This can be easily solved with brute force but will be improved in Section \ref{sec_inequality}. A further seven require negation of a specific fact which can be achieved with reasonable effort. The last ten are not deemed translatable because they require a feature that cannot be reasonably emulated in datalog. The feature mostly needed is a general form of negation that would allow to test for the non-existence of a derived fact. Other features needed for these constraints are regular expressions, which could theoretically be translated but would require large amounts of pre-processing, value comparison, and finally web requests to test if a link is valid.

@@

A problem that occurs multiple times is the need to determine inequality between constants. \emph{distinct value}, \emph{allowed units}, \emph{allowed qualifiers}, \emph{one of}, \emph{single best value} and \emph{one-of qualifier value} each require a rule to determine if a value, unit or property is not equal to another. To this end a certain amount of preprocessing is necessary, most notably the extraction of all constant pairs whose inequality has to be determined for finding the violations. To avoid putting and undue amount of processing outside of datalog this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g., all values on statements constrained by \emph{distinct value} or all qualifiers on statements constrained by \emph{allowed qualifiers}. From this set inequality will be generated, different versions of which will be introduced in Section \ref{sec_inequality} and evaluated with regards to their performance in Section \ref{sec_inequality_evaluation}.\\

Negation as a general problem cannot be solved in datalog. One could pre-process every necessary fact to be negated and add this to the ground facts, but this is not the purpose of this work. However, not all constraints require a general negation. \emph{item requires statement}, \emph{value requires statement}, \emph{inverse }and \emph{symmetric} are all violated if a statement does not exist on a specific item. This problem can be solved with limited amounts of preprocessing by leveraging a preprocessed statement order per item and 'iterating' through the statements until the required conditions are met.\\
A similar approach can be taken to find items with exactly one statement matching certain criteria, required for 	\emph{multi-value}. The idea is to ask the question if a second statement exists, then evaluating it as above. Both methods will be explained in Section \ref{sec_statement_non-existence}.

@@

To evaluate the rules described in Chapter \ref{cha_wikidata_constraints_explained} they have been implemented for VLog4J \cite{vlog4j}. VLog4J is a java library base on VLog \cite{vlog}, a Datalog engine utilising a column-based memory layout to allow processing of large datasets on commodity hardware. It was first published in 2016\cite{UJK2016} but has been actively developed since then.

The entire content of Wikidata is available in several formats: the aforementioned RDF, JSON and XML. To evaluate the violations, Wikidata needs to be provided in form of atoms as described in 
Section \ref{subsec_representation}. For this purpose the Wikidata Toolkit (WDTK)\cite{wdtk} was used. It provides an iterator over the JSON dumps, allowing the processing of all entities in Wikidata.

@@

Based on this set the Inequaliy Helper creates all files necessary to established inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen.

The rules determining violations are created by the Constraint Checker, which uses the Wikidata Query Service\footnote{\url{query.wikidata.org}} to fetch the current constraint configurations. Based on these configurations the rules are created as described in Chapter \ref{cha_wikidata_constraints_explained}.

Computing all violations based on the provided (inequality)-facts and violations rules is done per constraint. All facts pertaining to the constraint as well as the rules are loaded into VLog, which creates and in-memory database from the provided .csv-files. Once this is done, the reasoning engine materializes all possible facts. Afterwards, all violations are extracted by querying the database for the queries described in Section \ref{subsec_task}.

@@

The experiments where carried out on a server with a 3.50GHz Intel Xeon CPU, 1 TB SSD and 32 GB assigned RAM running Debian 4.9. Multiple sets of sizes ranging from one thousand to one million entities where extracted from the Wikidata JSON-Dump and all constraints evaluated against them.

@@

Because of a bug in VLog\footnote{\url{https://github.com/knowsys/vlog4j/issues/43}}, the largest dataset that could be completely evaluated contained 100,000 entities. With datasets of larger size, more than half of the constraints were affected by this bug, meaning the analysis could not be conducted. All further statistics describe the results on the 100,000 entity set unless specifically noted otherwise. Since inequality could only be computed in demanded mode (see Section  \ref{sec_inequality_evaluation}), all statistics will show the results attained in this mode.

@@

\begin{figure}
\includegraphics[width=\linewidth]{images/rules.pdf}
\caption{Number of rules and demanded rules}\label{fig_rules}
\end{figure}

\subsection{Facts}
Figure \ref{fig_facts} shows the number of facts loaded into VLog per constraint. Base facts denote those directly extracted from the JSON-Dump and corresponding to one of the predicates listed in Section \ref{subsec_representation}. Inequality are the letter\(_i\) facts created to establish inequality as described in Section \ref{sec_constraints_using_inequality}. Note that in the demanded (and encoded) mode, every entry for which inequality should be established is usually split into multiple letter\(_i\) facts based on its length, which is why the number of inequality facts can be up to ten times the number of base facts. Require facts list the order of statements or qualifiers as described in Section \ref{sec_constraints_using_statement_non}.

@@

\subsection{Inequality evaluation}\label{sec_inequality_evaluation}
Inequality computation was implemented in three variants outlined in Section \ref{sec_constraints_using_inequality}: naive, encoded and demanded. Naive and encoded could not terminate for any dataset other than the smallest with 1,000 entities. Naive failed because the disk space needed for the precomputed inequality facts was not available. Encoded, while improving on the disk space needed, still failed due to out of memory errors. This behaviour is expected, since for an input of size \(n\) \(n^2\) inequalities have to be materialized. The runtimes for this dataset, presented in Figure \ref{fig_inequality}, shows a clear improvement when using the demanded approach. This shows that the overhead in rules introduced to enable the demanded approach, seen in Figure \ref{fig_rules}, is worth the improvements in performance gained through it.

\begin{figure}
\includegraphics[width=\linewidth]{images/inequality1000.pdf}
\caption{Runtime materialization for 1,000 entities (encoded and demanded)}\label{fig_inequality}
\end{figure}

@@

\subsection{Distinct values}
While the rules for the \emph{distinct values} constraint (see Section \ref{subsec_distinct_values}) are relatively simple in writing, they do require a large amount of processing. Distinct values requires the comparison of every constrained statement with every other constrained statement, meaning its evaluation would scale badly due to the quadratic growth of statements to be compared. Additionally, the number of inequalities that have to be evaluated according to the demanded approach (see Section \ref{sec_constraints_using_statement_non}) grows accordingly. Due to the nature of the constraint -- requiring distinct values for a property across all of Wikidata -- this is in all likelihood unavoidable.

>>>>

While this is similar to RDF and Wikidata has in fact been exported as RDF \cite{EGKMV2014}, the data structure is more complex than simple triples. It allows to store additional information about statements. This is done via a statement ID, which is assigned to every triple storing information about an item. Statements have one of three ranks, \emph{normal}, \emph{preferred}, and \emph{deprecated} which are used to differentiate similar statements on the same item. For example, the \emph{preferred} rank would be given to the current \emph{head of state} (P35) of a country, while all statements denoting former \emph{heads of state} have the rank \emph{normal}. \emph{Deprecated} is used on statements that where once considered correct but are not anymore, e.g., \emph{Chelsea Manning} (Q298423) having the \emph{sex or gender} (P21) \emph{male} (Q6581097).

Every statement can have added qualifiers, which consist of the corresponding statement id, a predicate, and an object. Qualifiers provide contextual information for statements. The most common context is the time frame in which the statement was accurate, e.g., stating the \emph{start time} (P580) and \emph{end time} (P582) of the term of a head of state, but other information is possible, e.g. determination method for measurements or the elevation above sea level for the highest point. Additionally, references can be added to statements.

References support the claim by citing the source and additional information, e.g. when it was retrieved or when the source was published as property-value pairs. Each reference consists of multiple such pairs, and statements can have multiple references.

@@

\section{Datalog}
Datalog is a logic query language. It is fully declarative and uses rules to derive additional knowledge from given facts. The most basic domain element in datalog is a \emph{constant}, which will be denoted as \emph{someConstant}. In the context of Wikidata, any item or property, e.g. \emph{Earth} (Q2) or \emph{head of government} (P6), could be represented as a constant. Relations in datalog are expressed using \emph{predicates} with a specified arity \(n\), written as predicate\([n]\). To stay close to Wikidata in our examples, let us take the predicate triple\([3]\) as an example. This would be used to express the \emph{item}-\emph{property}-\emph{value} relation with each position representing one part of the triple.\\
Additionally, datalog also provides \emph{variables}, which will be capitalized, like \(X\). The sets of predicates {\bf P}, constants {\bf C} and variables {\bf V} are mutually disjoint. Constants and variables are collectively known as \emph{terms} and are, together with predicates, used to form \emph{atoms}. An atom has the form predicate\((t_1, \ldots , t_n)\) with \(n\) as the arity of the predicate and \(t_1,\ldots,t_n\) terms.\\
Atoms are the building blocks of \emph{rules}, which are used to derive information in datalog. A rule is a premise that implies a conclusion, and both are conjunctions of atoms.
Let \(P_n\) denote  premise atoms and \(C_n\) the conclusion atoms. The variables used in one rule are divided into three mutually disjoint sets \(V_p\), \(V_c\), and \(V_r\) which are the variables occurring only in its premise atoms, only in its conclusion atoms, and occurring in both, respectively. The general from of an existential rule is (with \(k \in \mathbb{N}^+\), \(l \in \mathbb{N}\))
\begin{equation}
\forall V_p \forall V_r.(\exists V_c.C_1 \wedge \ldots \wedge C_k \leftarrow P_1 \wedge \ldots \wedge P_l)
\end{equation}

\(V_c\) are called existential variables. If \(V_c = \emptyset\) the rule is called a \emph{datalog rule} and if \(l = 0\) it is called a \emph{fact}. If a fact contains no variables but only constants, it is a \emph{ground rule}. Since all \(V_p\) and \(V_r\) are  quantified, the preceding universal quantifiers \(\forall V_p \forall V_r\) will be omitted for the rules presented in this work.

@@

\begin{example}
For the following example, taken from Wikidata, descriptive names will be used instead of the IDs to increase readability. The individual entities and properties are introduced with their ID beforehand, so the example can be retraced. The species \emph{blue whale} (Q42196) has the \emph{parent taxon} (P171) \emph{balaenoptera} (Q133320), which in turn has the parent taxon \emph{rorqual} (Q232829). The predicates used are triple\_EDB\([3]\) as the input and isInTaxon\_IDB\([2]\), which will be derived. The basic facts are:
\[\text{triple\_EDB}(\text{blue whale}, \text{parent taxon}, \text{balaenoptera})\]
\[\text{triple\_EDB}(\text{balaenoptera}, \text{parent taxon}, \text{rorqual})\]
From this we can draw conclusions. The first step is to represent the already existing facts as IDB:
\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{triple\_EDB}(X, \text{parent taxon}, Y)\]
Finally the transitive closure of this relation can be derived with
\[\text{isInTaxon\_IDB}(X, Y) \leftarrow \text{isInTaxon\_IDB}(X, Z) \wedge \text{isInTaxon\_IDB}(Z, Y)\]
\end{example}

@@

To improve the data quality, to help editors avoid common mistakes, and to clarify the usage of properties, Wikidata allows constraints to be defined on properties. They are, however, not firm rules, and exceptions can and should be made if necessary. A property constraint says something about the way a property may or may not be used. This can mean simple restrictions on the statement a property is part of, e.g. that only certain values are allowed, or more complex requirements on connected parts of the data, e.g. that the value of a statement with this property should have certain statements. In Table \ref{tab_property_constraints} you can find the existing property constraints as of September 2018. Afterwards, the constraints will be explained in detail and their modelling given. For constraint that cannot be modelled, reasoning for this assessment is included.

To apply a constraint to a property, a statement <property> <P2302> <constraint> is added to the properties item page, where <property> is the property that should be constrained, <P2302> (property constraint) denotes that the property has a constraint and <constraint> is the specific constraint. Note that a constrained statement means a statement whose predicate is a property that has a specific constraint. A simple example for this would be \emph{distinct values} (Q21502410), stating that values of the property must be unique across all constrained statements. For this case no additional information regarding the constraint is necessary and violation could be found by surveying all constrained statements and comparing their values.\\
Constraints can be further specified by using qualifiers on the constraint statement. These qualifiers are either optional or mandatory, and may be applied once or multiple times. For example, the \emph{none of} constraint (Q52558054) states that values for the specific property may not be in a given set of disallowed values. This set is specified using qualifiers with property \emph{item of} (P2305), which list all the disallowed values. At least one such qualifier must be present for the statement.\\
More complex dependencies are also possible. A constraint can depend on all constrained statements of a specific item (\emph{single value} constraint [Q19474404]), on all statements of the item specified as the value of a constrained statement (\emph{value requires statement} constraint [Q21510864]), or on all statements using the subclass of-predicate (\emph{type} constraint [Q21503250]).

Table \ref{tab_property_constraints} shows all constraints that are used as of September 2018. It is roughly sorted by how difficult it is to model the given constraint using logical rules, based on the required expressive features and pre-processing. These required features are conditions that cannot natively be expressed in datalog and would require some kind of pre-processing of the dataset. Four constraints do not need any pre-processing except for loading the data into datalog as ground rules. Six can be solved if constant inequality is defined. This can be easily solved with brute force but will be improved in Section \ref{sec_constraints_using_inequality}. A further seven require negation of a specific fact which can be achieved with reasonable effort. The last ten are not deemed translatable because they require a feature that cannot be reasonably emulated in datalog. The feature mostly needed is a general form of negation that would allow to test for the non-existence of a derived fact. Other features needed for these constraints are regular expressions, which could theoretically be translated, but would require large amounts of pre-processing, value comparison, and finally web requests to test if a link is valid.

@@

A problem that occurs multiple times is the need to determine inequality between constants. \emph{distinct value}, \emph{allowed units}, \emph{allowed qualifiers}, \emph{one of}, \emph{single best value} and \emph{one-of qualifier value} each require a rule to determine if a value, unit, or property is not equal to another. To this end a certain amount of preprocessing is necessary, most notably the extraction of all constant pairs whose inequality has to be determined for finding the violations. To avoid putting and undue amount of processing outside of datalog, this extraction will not check more than that the constant is in the right position relative to the constrained statement, e.g., all values on statements constrained by \emph{distinct value} or all qualifiers on statements constrained by \emph{allowed qualifiers}. From this set inequality will be generated, different versions of which will be introduced in Section \ref{sec_inequality} and evaluated with regards to their performance in Section \ref{sec_constraints_using_inequality}.\\

Negation as a general problem cannot be solved in datalog. One could pre-process every necessary fact to be negated and add this to the ground facts, but this is not the purpose of this work. However, not all constraints require a general negation. \emph{item requires statement}, \emph{value requires statement}, \emph{inverse }and \emph{symmetric} are all violated if a statement does not exist on a specific item. This problem can be solved with limited amounts of preprocessing by leveraging a preprocessed statement order per item and 'iterating' through the statements until the required conditions are met.\\
A similar approach can be taken to find items with exactly one statement matching certain criteria, required for 	\emph{multi-value}. The idea is to ask the question if a second statement exists, then evaluating it as above. Both methods will be explained in Section \ref{sec_constraints_using_statement_non}.

@@

To evaluate the rules described in Chapter \ref{cha_wikidata_constraints_explained} they have been implemented for VLog4J \cite{vlog4j}. VLog4J is a Java library based on VLog \cite{vlog}, a Datalog engine utilising a column-based memory layout to allow processing of large datasets on commodity hardware. It was first published in 2016\cite{UJK2016} but has been actively developed since then.

The entire content of Wikidata is available in several formats: RDF (as mentioned), JSON and XML. To evaluate the violations, Wikidata needs to be provided in form of atoms as described in 
Section \ref{subsec_representation}. For this purpose the Wikidata Toolkit (WDTK)\cite{wdtk} was used. It provides an iterator over the JSON dumps, allowing the processing of all entities in Wikidata.

@@

At the same time the Inequaliy Helper, receiving the Facts from WDTK, creates all files necessary to establish inequality as described in Section \ref{sec_constraints_using_inequality}, based on which approach was chosen.

The rules determining violations are created by the Constraint Checker, which uses the Wikidata Query Service (WDQS)\footnote{\url{query.wikidata.org}} to fetch the current constraint configurations, which are the set constraints and their qualifiers as described in Chapter \ref{cha_property_constraints}. Based on these configurations the rules are created as described in Chapter \ref{cha_wikidata_constraints_explained}.

Computing all violations based on the provided (inequality)-facts and violations rules is done per constraint type. All facts pertaining to the constraint as well as the rules are loaded into VLog, which creates and in-memory database from the provided .csv-files. Once this is done, the reasoning engine materializes all possible facts. Afterwards, all violations are extracted by querying the database for the queries described in Section \ref{subsec_task}.

@@

The experiments where carried out on a server with a 3.50GHz Intel Xeon CPU, 1 TB SSD and 32 GB assigned RAM running Debian 4.9. Multiple sets of sizes ranging from one thousand to one million entities where randomly chosen from the Wikidata JSON-Dump and all constraints evaluated against them.

@@

Because of a bug in VLog\footnote{\url{https://github.com/knowsys/vlog4j/issues/43}}, the largest dataset that could be completely evaluated contained 100,000 entities. With datasets of larger size, more than half of the constraints were affected by this bug, meaning the analysis could not be conducted. All further statistics describe the results on the 100,000 entity set unless specifically noted otherwise. Since inequality could only be computed in on-demand mode (see Section  \ref{sec_inequality_evaluation}), all statistics will show the results attained in this mode.

@@

\begin{figure}
\includegraphics[width=\linewidth]{images/rules.pdf}
\caption{Number of rules and on-demand rules}\label{fig_rules}
\end{figure}

\subsection{Facts}
Figure \ref{fig_facts} shows the number of facts loaded into VLog per constraint. Base facts denote those directly extracted from the JSON-Dump and corresponding to one of the predicates listed in Section \ref{subsec_representation}. Inequality are the letter\(_i\) facts created to establish inequality as described in Section \ref{sec_constraints_using_inequality}. Note that in the on-demand (and encoded) mode, every entry for which inequality should be established is usually split into multiple letter\(_i\) facts based on its length, which is why the number of inequality facts can be up to ten times the number of base facts. Require facts list the order of statements or qualifiers as described in Section \ref{sec_constraints_using_statement_non}.

@@

\subsection{Inequality evaluation}\label{sec_inequality_evaluation}
Inequality computation was implemented in three variants outlined in Section \ref{sec_constraints_using_inequality}: naive, encoded and on-demand. Naive and encoded could not terminate for any dataset other than the smallest with 1,000 entities. Naive failed because the disk space needed for the precomputed inequality facts was not available. Encoded, while improving on the disk space needed, still failed due to out of memory errors. This behaviour is expected, since for an input of size \(n\), the number of inequalities that have to be materialized is \(n^2\). The runtimes for this dataset, presented in Figure \ref{fig_inequality}, shows a clear improvement when using the on-demand approach. This shows that the overhead in rules introduced to enable the on-demand approach, seen in Figure \ref{fig_rules}, is worth the improvements in performance gained through it.

\begin{figure}
\includegraphics[width=\linewidth]{images/inequality1000.pdf}
\caption{Runtime materialization for 1,000 entities (encoded and on-demand)}\label{fig_inequality}
\end{figure}

@@

\subsection{Distinct values}
While the rules for the \emph{distinct values} constraint (see Section \ref{subsec_distinct_values}) are relatively simple in writing, they do require a large amount of processing. Distinct values requires the comparison of every constrained statement with every other constrained statement, meaning its evaluation would scale badly due to the quadratic growth of statements to be compared. Additionally, the number of inequalities that have to be evaluated according to the on-demand approach (see Section \ref{sec_constraints_using_statement_non}) grows accordingly. Due to the nature of the constraint -- requiring distinct values for a property across all of Wikidata -- this is in all likelihood unavoidable.
